---
title: "Tidy data and relational datasets"
author: "Yiying Wu"
date: "2023-09-26"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
options(tibble.print_min = 5)
```


## `pivot_longer`
import sas dataset
```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names()

pulse_df
```

pivot wide to long
```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    values_to = "bdi")

pulse_tidy_data
```

`names_prefix`: don’t keep the `bdi_score_` prefix 

```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi")

pulse_tidy_data
```

`replace`: changing `bl` to `00m`

converting `visit` to a factor variable

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)) 

print(pulse_df, n = 12)
```

**Learning Assessment:** In the litters data, the variables gd0_weight and gd18_weight give the weight of the mother mouse on gestational days 0 and 18. Write a data cleaning chain that retains only litter_number and these columns; produces new variables gd and weight; and makes gd a numeric variable taking values 0 and 18 (for the last part, you might want to use recode …). Is this version “tidy”?

`case_match`: replace gd0_weight with 0, replace gd18_weight with 18. Then change gd into numeric variable
```{r}
litters_wide = 
  read_csv("./data/FAS_litters.csv") |>
  janitor::clean_names() |>
  select(litter_number, ends_with("weight")) |> 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd", 
    values_to = "weight") |> 
  mutate(
    gd = case_match(
      gd,
      "gd0_weight"  ~ 0,
      "gd18_weight" ~ 18
    ))
```

## `pivot_wider`
```{r}
analysis_result = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )

analysis_result
```

```{r}
pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean")
```
use select to reorder columns, and (depending on your goal) use knitr::kable() to produce a nicer table for reading.


## `Binding rows`
`range` which column to read
```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")
```

